---
title: "Team Project by Xiaoming Ji"
author: "STAT 420, Summer 2017, Xiaoming Ji"
date: ''
output:
  html_document: 
    toc: yes
  pdf_document: default
---

We load the data and fit it with a simple regression model with SalePrice as response and LotArea as predictor.


```{r, message=FALSE}
library(readr)
house_data = read_csv("HousePrices/train.csv")

price_model = lm(SalePrice ~ LotArea, data = house_data)
```

We draw a plot to show the data relation and the fitted line.
```{r, echo=FALSE}
plot(SalePrice ~ LotArea, data = house_data, col = "grey", pch = 20,
     main = "Price by LotArea")
abline(price_model, col = "darkorange", lwd = 3)
```

The plot shows the data is gathers in the low range of LotArea. Let's try a log transforming the predictor and replot.

```{r}
price_model_log = lm(SalePrice ~ log(LotArea), data = house_data)
```

```{r, echo=FALSE}
plot(SalePrice ~ log(LotArea), data = house_data, col = "grey", pch = 20,
     main = "SalePrice by Log(LotArea)")
abline(price_model_log, col = "darkorange", lwd = 3)
```

The new plot seems much better. 

We use the corrplot package in combination with the cor() function to plot the correlations between data’s selected attributes.

```{r, echo=FALSE}
library(corrplot)
colunms = c("SalePrice", "LotArea", "MSSubClass", "MSZoning", "LotConfig")
house_data_reg = with(house_data, data.frame(SalePrice, LotArea, 
                                             MSSubClass = as.numeric(as.factor(MSSubClass)),
                                             MSZoning = as.numeric(as.factor(MSZoning)), 
                                             LotConfig = as.numeric(as.factor(LotConfig)),
                                             Neighborhood = as.numeric(as.factor(Neighborhood)), 
                                             BldgType = as.numeric(as.factor(BldgType)),
                                             HouseStyle = as.numeric(as.factor(HouseStyle)),
                                             OverallQual = as.numeric(as.factor(OverallQual)),
                                             OverallCond = as.numeric(as.factor(OverallCond)),
                                             YearBuilt, YearRemodAdd,
                                             RoofStyle = as.numeric(as.factor(RoofStyle)),
                                             ExterQual = as.numeric(as.factor(ExterQual)),
                                             ExterCond = as.numeric(as.factor(ExterCond)),
                                             TotalBsmtSF, FullBath, BedroomAbvGr, 
                                             KitchenAbvGr, Fireplaces, GarageCars))

corrplot(cor(house_data_reg), method="circle")
```

The covariance test shows `OverallQual`, `TotalBsmtSF`, `YearBuilt`, `YearRemodAdd`, `ExterQual`, `FullBath`, `Fireplaces` has high relation with `SalePrice`.

```{r}
library(faraway)
price_model_reg = lm(SalePrice ~ OverallQual + TotalBsmtSF + YearBuilt + 
                       YearRemodAdd + ExterQual + FullBath + Fireplaces,
                     data = house_data_reg)
vif(price_model_reg)
```

